---
title: "Exercises for git and GitHub"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
# Exercises
1. Go to GitHub and create an account, if you have not yet done so.
2. Run the file `install_R_packages_gitwebinar.R` from the `/setup` folder,
which will check that you have git installed on your system, and will download all of the required **R** packages.

## git_for_r presentation
3. Fork the GitHub repository `CentreForHydrology/git_for_R.git` to your account. 

Note that you could clone directly from `https://github.com/CentreForHydrology/git_for_R.git`, but it is better practice for collaboration to work from your own copy of the repo.

4. Make a local copy of the presentation repository by cloning your repository.  
(`https://github.com/<username>/git_for_r`).

```{r checkout_git_for_r, eval=FALSE, include=FALSE}
username <- "rchlumsk" # github username
useremail <- "rchlumsk@gmail.com"

setwd("C:/Users/rober/Documents/GitHub/")
system(sprintf("git clone https://github.com/%s/git_for_R", username))
setwd("git_for_R/")
```

5. Delete the file `this_file_is_junk.txt` from the local repo. It is inside the `/exercises` folder.

6. Make a change to the `README.md` file in the main folder, such as adding a fun note to this file.

7. Commit the changes to your local repo; be sure to stage the changes first.

```{r commit_git_for_r, eval=FALSE, include=FALSE}
# check the status
system("git status")

# add the unstaged changes (add all changes)
system("git add -A")

## configure username if needed
system(sprintf("git config --global user.name %s", username))
system(sprintf("git config --global user.email %s", useremail))

# commit the changes
system("git commit -m \"removed exercises/this_file_is_junk.txt, updated README.md document\"")
```

8. Push the changes to your local repo. Check that the README.md page on your github repo online has been updated.

```{r push_git_for_r, eval=FALSE, include=FALSE}
system("git push origin master")
```

9. At this stage, if you were happy with your changes, you might submit a pull request. This would 'request' the managers of the original repo to consider pulling your updates into their repo.

## CSHS-hydRology package
10. Fork the GitHub repository `CSHS-CWRA/CSHShydRology` to your account.

11. Create a local version of the `CSHShydRology` package on your own computer by creating a new project in RStudio. 

```{r checkout_hydRology, eval=FALSE, include=FALSE}
username <- "rchlumsk" # github username
useremail <- "rchlumsk@gmail.com"

setwd("C:/Users/rober/Documents/GitHub/")
system(sprintf("git clone https://github.com/%s/CSHShydRology", username))
setwd("CSHShydRology/")
```

12. Build the package. Check to see that it builds properly. For more in-depth checking, use the **Check** command.

13. Create a branch called `testing` for the CSHShydRology repo on your computer. You may need to revert changes made to files before switching or creating branches.

14. Create a file in the /R folder called "this_is_junk.txt" with a fun message.

15. Commit the changes to the local repo and push the commit to GitHub. Do this without building the package for now (noting that this is in line with bad practices).

```{r commit_and_push_hydRology, eval=FALSE, include=FALSE}
# check the status
system("git status")

# add the unstaged changes (add all changes)
system("git add -A")

## configure username if needed
system(sprintf("git config --global user.name %s", username))
system(sprintf("git config --global user.email %s", useremail))

# commit the changes
system("git commit -m \"added junk file to R folder\"")

# push the changes
system("git push origin testing")
```

16. Check the `testing` branch. Does the package build properly? Change the junk file so that the testing branch will build.

17. Change back to the `DanMoore_spatial` branch and note the differences in files in the /R folder.